---
knit: (function(input, ...) {
    rmarkdown::render(input, output_dir = "03-render-output")
  })
output:
  html_document:
    df_print: paged
  html_notebook: default
---

#### Setup
```{r setup-configuration, echo = F, message = F, warning = F, results = 'hide'}
knitr::opts_chunk$set(echo = F, results = "asis")

# USE_GCS <- TRUE ## use this if running scan directly from folder and steam data from gcs
source("R/setup.R")
source("R/fns.R")
source("https://raw.githubusercontent.com/compoundrisk/monitor/databricks/src/fns/helpers.R")

aoi <- aoi %>% st_as_sf() %>% st_transform("EPSG:4326")


# Disable s2 geometry checks to handle invalid geometries
sf::sf_use_s2(FALSE)

# country_groups <- read_csv("https://raw.githubusercontent.com/compoundrisk/monitor/databricks/src/country-groups.csv")  # not sure if this is necessary



# Change the print type depending on whether document is run interactively or from `rmarkdown::render()`
# This is because a single function (such as `if(in_oxford)` can only print one result)
# from_render is defined in render-current-scans.R
# when rendering, must also set results to "asis" if using print_paged_df

interactive_run <- if (exists("from_render")) F else T
results_type <- if (!interactive_run) "asis" else "markup"

# helper
print_paged_df <- function(...) {
  cat(rmarkdown:::print.paged_df(rmarkdown::paged_table(...)))
}

print_text <- function(x, linebreaks = 2) {
  cat(paste0(x, "\n", paste(rep("<br>", linebreaks), collapse = ""), "\n"))
}


# ```{r, echo = F}
# This chunk is separate from the setup chunk above because we don't want
# the message suppressed 

# c("03-render-output/plots", "data-sharing", "data-sharing/02-TIFs",
#   "data-sharing/03-analysis-inputs", "data-sharing/03-analysis-inputs/flood-events",
#   "data-sharing/03-analysis-inputs/earthquakes", "data-sharing/04-tabular-data") %>%
#   lapply(function(directory) {
#     if (!dir.exists(directory)) dir.create(directory) else {
#       # message(paste0(directory, "/ directory already exists in ", getwd()))
# }
#   }) %>% invisible()

# if (!dir.exists("data-sharing/AOI")) {
# file.copy(from = "01-user-input/AOI/", to = "data-sharing", recursive = T)
# file.rename("data-sharing/AOI", "data-sharing/01-AOI")
# }

# message("To fill data-sharing/02-TIFs and data-sharing/03-analysis-inputs you will need to download and rename files from city_scan2. This is not yet a necessary step for all cities.")
```

```{r pre-charting, echo = F, message= F, warning=F, results = 'hide'}  
source("R/pre-charting.R")
```

##### _User Input Variables_
```{r setting-title, echo = F}

print_text(paste("scan-ID", paste_bold(scan_id), ' | GCS: ', paste_bold(USE_GCS)), 2 )

print_text(paste("City:", paste_bold(city)), 1)
print_text(paste("Country:", paste_bold(country)), 1)
if (length(bm_cities_manual) == 0) {
  print_text(paste("Manual benchmark cities:", paste_bold("None defined - using automatic detection from nearby countries")), 1)
} else {
  print_text(paste("Manual benchmark cities:", paste_bold(paste_and(bm_cities_manual))), 1)
}

```

---
title: "`r city` City Scan"
---

<!-- ### Rename maps & make sure all are included -->


### Area and Climate Zone

```{r koeppen-classification, echo = F}
# Basic Info (Area, Climate)
# Area
aoi <-  aoi %>% st_as_sf() %>% st_make_valid() # change to sf object (easiest)
area <- units::set_units(st_area(st_combine(aoi)), km^2)
# Get area
print_text(paste("AOI Area:", paste_bold(round(area, 1)), "sq km"), 1)

# Climate
coords <- st_centroid(aoi) %>% st_coordinates() %>% suppressWarnings()
names(coords) <- c("Lon", "Lat")       
koeppen <- read_csv(koeppen_file, col_types = "ddc")
koeppen_city <- subset(koeppen, between(Lon, coords["Lon"] - .5, coords["Lon"] + .5) &  
                         between(Lat, coords["Lat"] - .5, coords["Lat"] + .5)) %>%
  .$Cls %>% unique()
# Find class here: https://en.wikipedia.org/wiki/Köppen_climate_classification

print_text(paste("Köppen climate classification:", paste_bold(koeppen_city), "(See https://en.wikipedia.org/wiki/Köppen_climate_classification for classes)"), 1)

```


### Population Data Setup
```{r population-data-setup, echo = F, message= F, warning=F}

#  Check if city is in Oxford Economics Dataset 

if (in_oxford) {

  print_text(paste0(paste_bold(city), " is in **Oxford Economics**.", " Total Population: ", paste_bold(scales::label_comma()(pop * 1000)) ), 1)

  } else {
  
  print_text(paste0(paste_bold(city), " is not in Oxford Economics. The included locations for ", 
            paste_bold(country), " are ",
            paste_bold(paste_and(oxford_locations_in_country$Location))
                    # ". Is one of these an alternate name or spelling for ", city, "?"
                    ))

  print_text(paste( "Using", paste_bold(pop_source), "for",  paste_bold(city), "Total Population:", paste_bold(scales::label_comma()(pop * 1000)) ) )

}

if (nrow(wpop_growth) > 0 && "Year" %in% names(wpop_growth)){
  print_text(paste0("WorldPop data projects population of ", paste_bold(city), " grew from ", 
            paste_bold(scales::label_comma()(wpop_growth$Population[wpop_growth$Year == min(wpop_growth$Year)])),
            " in ", paste_bold(min(wpop_growth$Year)), " to ",
            paste_bold(scales::label_comma()(wpop_growth$Population[wpop_growth$Year == max(wpop_growth$Year)])),
            " in ", paste_bold(max(wpop_growth$Year)), ".") ,1)
}

# here, `pop` is the total population for the scan city
```



### Benchmark Cities
Benchmark cities are selected as cities that are in neighboring countries and having a population of 50% to 150% of the focus city's population, with addtional cities manually specified.

```{r benchmark-cities-selection, echo = F}
# Read Data ----

# Print table of population in benchmark cities (OUTPUT)
print_text("Benchmark cities")

pop_longitude %>%
    subset(Location %in% bm_cities) %>%
    group_by(Location) %>%
    slice_min(abs(Year - 2021), n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    arrange(Location)

# Print table of population of nearby cities not in benchmark list (OUTPUT)
print_text("Nearby non-benchmark cities")
oxford_full %>%
  select(Location, Country, Indicator, `2021`) %>%
  subset(Location %in% nearby_cities & Location %ni% bm_cities & Indicator == "Total population") %>%
  mutate(`2021` = `2021` * 1000) %>%
  arrange(`2021`) 

```

```{r oxford-data-setup, echo = F}
# Subset ofxfor data for city and benchmark cities
oxford <- subset(oxford_full, Location %in% c(city, bm_cities)) %>%
  mutate(Group = case_when(Location == city ~ Location, T ~ "Benchmark") %>%
           factor(levels = c(city, "Benchmark")))


# Generate benchmark city text (OUTPUT)
print_text(paste_bold("Benchmark footnote text:"))

print_text("Following cities are included in Oxford Economics dataset:", 1)
oxford %>% 
  select(Location, Country) %>% distinct() %>%
  arrange(Country) %>% group_by(Country) %>%
  summarize(text = paste_and(Location), .groups = "drop") %>%
  mutate(text = paste0(text, " (", Country,")")) %>%
  .$text %>% paste_and() %>%
  paste_bold() %>%
  print_text()


```




### Population Growth
```{r population-growth, echo = F, message= F, warning=F}

# use worldpop data if available
if (exists("wpop_growth") && nrow(wpop_growth) > 0 && "Year" %in% names(wpop_growth)) {
  pop_growth <- wpop_growth
}

# Check if pop_growth exists and has data
if (!exists("pop_growth") || is.null(pop_growth) || nrow(pop_growth) == 0) {
  print_text(paste_bold("No population growth data available"))
} else {

# Make pop_growth plot
pop_min_year <- min(pop_growth$Year)
pop_max_year <- max(pop_growth$Year)

pop_growth_plot <- ggplot(pop_growth, aes(x = Year, y = Population, group = Location)) + #, color = Source))
  geom_line() +
  geom_point() +  
 
  scale_x_continuous(
    breaks = if (pop_max_year - pop_min_year > 15)
      seq(floor(pop_min_year/5)*5, ceiling(pop_max_year/5)*5, by = 5)
    else
      seq(floor(pop_min_year/2)*2, ceiling(pop_max_year/2)*2, by = 2),
    minor_breaks = seq(pop_min_year, pop_max_year, 1)) + 
  scale_y_continuous(limits = c(0, max(pop_growth$Population)), labels = scales::label_comma(scale_cut = scales::cut_short_scale()), expand = expansion(c(0, .1))) +    
  theme_minimal() +
  labs(title = paste0(city, " Population Growth, ", pop_min_year, "-", pop_max_year)) +
  theme(
    axis.line = element_line(linewidth = .5, color = "black"),
    panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
    panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray"))
  # For when we want to plot multiple components to the city population
  # scale_color_manual(values = c(Combined = "black", Other = "dark grey")) +
  # ggrepel::geom_text_repel(
  #   data = slice_max(pop_growth, by = Location, order_by = Year), aes(label = Location),
  #   segment.color = NA, hjust = 0, nudge_x = .5, direction = "y") +
  # theme(legend.position = "none")
ggsave("03-render-output/plots/oxford-pop-growth.png", plot = pop_growth_plot, device = "png",
       width = 8, height = 5, units = "in", dpi = "print")

# Print pop growth plot (OUTPUT)
pop_growth_plot

# The tibble and pmap approach is in case pop_growth includes multiple components
pop_growth_summary <- pop_growth %>%
  arrange(Year) %>%
  summarize(.by = Location, 
    year_1 = head(Year, 1),
    pop_1 = head(Population, 1),
    year_last = tail(Year, 1),
    pop_last = tail(Population, 1),
    span = year_last - year_1,
    change = pop_last - pop_1,
    change_pct = scales::percent(pop_last/pop_1 - 1, accuracy = 0.1),
    growth_rate = scales::percent((pop_last/pop_1)^(1/span) - 1,  accuracy = 0.1))
print_paged_df(pop_growth_summary)
pmap(pop_growth_summary, \(...) {
  row <- tibble(...)
  glue("{paste_bold(row$Location)}'s population grew by {paste_bold(row$change_pct)} from {paste_bold(scales::comma(row$pop_1))} in {paste_bold(row$year_1)} to {paste_bold(scales::comma(row$pop_last))} in {paste_bold(row$year_last)} for an average annual growth rate of {paste_bold(scales::percent(((row$pop_last/row$pop_1)^(1/(row$year_last-row$year_1)) - 1), accuracy = 0.1))}")
}) %>% unlist() %>% cat(sep = "\n") %>% print_text(1)

# Print pop growth tables (OUTPUT)
print_text(paste0(paste_bold("Yearly population"), " - sorted by absolute growth"))
pop_growth %>% mutate(growth = Population - lag(Population)) %>%
  arrange(Location, desc(growth)) 

print_text(paste0(paste_bold("Yearly population"), " - sorted by percentage growth"))
pop_growth %>% mutate(growth_pct = (Population - lag(Population))/lag(Population)) %>%
  arrange(Location, desc(growth_pct)) 
# if (in_oxford) {
#   print(paste(
#     "Average growth rate 2000-2010:",
#     pop_growth %>% mutate(growth_pct = (Population - lag(Population))/lag(Population)) %>%
#       filter(Year %>% between(2000, 2010)) %>%
#       arrange(desc(growth_pct)) %>% .$growth_pct %>% mean(na.rm = T) %>%
#       scales::percent(accuracy = 0.1)
#   ))
# }

} # end else for pop_growth check
```


### Population Density
<!-- Density should be overlaid histograms or box plots? joy division? When we have this spatial data ...  -->
```{r population-density, echo = F, message= F, warning=F}

hues <- c("black", "dark grey") %>% #c(scales::hue_pal()(2)) %>% 
  setNames(c(city, "Benchmark"))
hues[1] <- "#000000"

density_plot <- density %>%
  filter(!is.na(Density)) %>%
  ggplot() +
  geom_point(aes(x = Population, y = Density, color = Group)) +
  geom_text_repel(
    aes(x = Population, y = Density, label = Location, color = Group),
    box.padding = 0.25, min.segment.length = 0.001, point.padding = 1,
    segment.linetype = "dotted") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.line = element_line(linewidth = .5, color = "black"),
    panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
    panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray")) + 
  labs(title = paste("Population density of", city, "and benchmark cities"),
       y = bquote('Population density, people per '~km^2),
       x = "Population") +
  scale_y_continuous(labels = scales::label_comma(),
                     expand = expansion(mult = c(0, .03))) +
  scale_x_continuous(labels = scales::label_comma()) +
  scale_color_manual(values = hues) +
  expand_limits(y = 0)

ggsave("03-render-output/plots/oxford-pop-density-scatter.png", plot = density_plot, device = "png",
       width = 8, height = 5, units = "in", dpi = "print")

# Print denstiy plot and table (OUTPUT)
density_plot

density %>% arrange(-Density)


```

### Population Distribution

#### Population distribution by Oxford Economics
`r if (!in_oxford) "_Only included if city is in Oxford Economics database. City is not._"`

```{r population-distribution-oxford, echo = F, message= F, warning=F}

if (in_oxford && exists("pop_dist_group")) {
  
  # Print population distribution (OUTPUT)
  print_paged_df(pop_dist_group)
  
  pop_distro_plot <- ggplot(pop_dist_group, aes(x = Age_Bracket, y = Percentage)) +
    # geom_bar(stat = "identity", position = "dodge2") + 
    geom_col(aes(fill = Group), position = "dodge") + 
    labs(title = paste("Population distribution in", city, "and benchmark cities"),
         y = "Percentage", x = "Age Bracket", fill = "") +
    theme_minimal() +
    theme(legend.position = "bottom",
          legend.key.height = unit(1/8, "in"),
          axis.line = element_line(linewidth = .5, color = "black"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(linewidth = .125, color = "dark gray"),
          panel.grid.minor.y = element_line(linewidth = .125, linetype = 2, color = "dark gray")
          # panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
          # panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray")
    ) + 
    scale_y_continuous(breaks = seq(0, .2, .01),
                       labels = scales::percent_format(accuracy = 1),
                       limits = c(0, ceiling(max(pop_dist_group$Percentage) * 100) /100),
                       expand = c(0,0)) +
    scale_fill_manual(values = hues)
  # limits = c(0,.1))
  ggsave("03-render-output/plots/oxford-pop-distro.png", plot = pop_distro_plot, device = "png",
         width = 8, height = 6.5, units = "in", dpi = "print")
  
  # Print population distribution plot (OUTPUT)
  ggplot2:::print.ggplot(pop_distro_plot)
  
  # # Plotting histogram of focus city and lines for all other cities
  # # Probably better to use just one line for the median city
  # pop_dist_long %>%
  #   ggplot(aes(x = Age_Bracket, y = Percentage, group = Location)) +
  #   # geom_col(aes(fill = Location), position = "dodge")
  #   geom_col(data = pop_dist_long %>% filter(Location == city)) +
  #   geom_line(color = "dark grey") +
  #  labs(title = paste("Population distribution in", city, "and benchmark cities"),
  #        y = "Percentage", x = "Age Bracket", fill = "") +
  #   theme_minimal() +
  #   theme(legend.position = "bottom",
  #         legend.key.height = unit(1/8, "in"),
  #         axis.line = element_line(linewidth = .5, color = "black"),
  #         panel.grid.major.x = element_blank(),
  #         panel.grid.minor.x = element_blank(),
  #         panel.grid.major.y = element_line(linewidth = .125, color = "dark gray"),
  #         panel.grid.minor.y = element_line(linewidth = .125, linetype = 2, color = "dark gray")
  #         # panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
  #         # panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray")
  #         ) + 
  #   scale_y_continuous(breaks = seq(0, .2, .01),
  #                      labels = scales::percent_format(accuracy = 1),
  #                      limits = c(0, ceiling(max(pop_dist_group$Percentage) * 100) /100),
  #                      expand = c(0,0))
}


```


#### Population by WorldPop

Reproductive age is percentage of female population that is of reproductive age

```{r worldpop-age-groups, echo = F, message= F, warning=F}
# WorldPop Age Groups

if (!exists('pop_dist_group_wp')) print_text(paste_bold("No WorldPop age distribution file available")) else {
  
  # Print WorldPop age-sex distribution table (OUTPUT)
  print_paged_df(pop_dist_group_wp)
  
  pop_age_sex_plot <- ggplot(pop_dist_group_wp, aes(x = Age_Bracket, y = Percentage, fill = Sex)) +
    # geom_bar(stat = "identity", position = "dodge2") + 
    geom_col(position = "dodge") + 
    labs(title = paste("Population distribution in", city, "by sex"),
         y = "Percentage", x = "Age Bracket", fill = "") +
    theme_minimal() +
    theme(legend.position = "none",
          legend.key.height = unit(1/8, "in"),
          axis.line = element_line(linewidth = .5, color = "black"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(linewidth = .125, color = "dark gray"),
          panel.grid.minor.y = element_line(linewidth = .125, linetype = 2, color = "dark gray")
          # panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
          # panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray")
    ) + 
    scale_y_continuous(breaks = seq(0, .2, .01),
                       labels = scales::percent_format(accuracy = 1),
                       limits = c(0, ceiling(max(pop_dist_group_wp$Percentage) * 100) /100),
                       expand = c(0,0)) #+
  # scale_fill_manual(values = hues)
  
  # Print WorldPop age-sex distribution plot (OUTPUT)
  ggplot2:::print.ggplot(pop_age_sex_plot)
  ggsave("03-render-output/plots/world-pop-age-sex.png", plot = pop_age_sex_plot, device = "png",
         width = 8, height = 6.84, units = "in", dpi = "print")
  
  under5 <- pop_dist_group_wp %>% 
    filter(Age_Bracket %in% c("0-4")) %>% summarize(across(c(Count, Percentage), sum))
  youth <- pop_dist_group_wp %>% 
    filter(Age_Bracket %in% c("15-19", "20-24")) %>% summarize(across(c(Count, Percentage), sum))
  working_age <- pop_dist_group_wp %>% 
    filter(Age_Bracket %in% c("15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64")) %>% summarize(across(c(Count, Percentage), sum))
  elderly <- pop_dist_group_wp %>% 
    filter(Age_Bracket %in% c("60-64", "65-69", "70-74", "75-79", "80+")) %>% summarize(across(c(Count, Percentage), sum))
  sex_totals <-  pop_dist_group_wp %>% 
    group_by(Sex) %>% summarize(across(c(Count, Percentage), sum))
  female_pct <- sex_totals[which(sex_totals$Sex == "Female"), 3]
  sex_ratio <- (1 - female_pct) / female_pct * 100
  reproductive_age <-  pop_dist_group_wp %>% 
    filter(Sex == "Female", Age_Bracket %in% c("15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49")) %>% summarize(across(c(Count, Sexed_Percent), sum))
  
  if (!in_oxford) {
    # Print age distribution if not in Oxford (OUTPUT)
    pop_dist_group_wp %>% group_by(Age_Bracket) %>% 
      summarize(Percentage = sum(Percentage)) %>%
      mutate(cumpct = cumsum(Percentage)) %>%
      print_paged_df()
  }
  
  # Print demographic stats (OUTPUT)
  print_text(paste("under5:", paste_bold(under5$Percentage %>% {scales::label_percent()(.)})))
  print_text(paste("youth (15-24):", paste_bold(youth$Percentage %>% {scales::label_percent()(.)})))
  print_text(paste("working_age (15-64):", paste_bold(working_age$Percentage %>% {scales::label_percent()(.)})))
  print_text(paste("elderly (60+):", paste_bold(elderly$Percentage %>% {scales::label_percent()(.)})))
  print_text(paste("reproductive_age, percent of women (15-50):", paste_bold(reproductive_age$Sexed_Percent %>% {scales::label_percent()(.)})))
  print_text(paste("sex_ratio:", paste_bold(round(sex_ratio, 2)), "males to 100 females"))
}


```

### Population Density Distribution
```{r population-ridge, echo = F, message= F, warning=F}

librarian::shelf(ggridges)

if (!exists("wpop_df")) {
  print_text(paste_bold("No population raster data available"))
} else{ 
  # Calculate x-axis limit (cap at 99th percentile)
  x_max <- quantile(wpop_df$pop, 0.99)

  # Create ridge plot with quantile coloring
  p_density <- wpop_df %>%
    ggplot(aes(x = pop, y = 1, fill = factor(after_stat(quantile)) )) +
    stat_density_ridges(
      geom = "density_ridges_gradient",
      calc_ecdf = TRUE,
      quantiles = 1:4/5,
      quantile_lines = FALSE,
      alpha = 0.6,
      scale = 0.9
    ) +
    scale_x_continuous(
      limits = c(0, x_max),
      breaks = seq(0, ceiling(x_max/10)*10, by = max(10, ceiling(x_max/100)*10)),
      minor_breaks = seq(0, ceiling(x_max/10)*10, by = max(5, ceiling(x_max/200)*10))
    ) +
    scale_fill_manual(
      name = "Percentile",
      values = c('#ECEB72', '#D4BE62', '#BD9252', '#A56542', '#8E3933'),
      labels = c("0-20th", "20-40th", "40-60th", "60-80th", "80-100th")
    ) +
    labs(x = "Population density (people/hectare)") +
    theme_minimal() +
    theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      legend.position = "bottom",
      panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
      panel.grid.minor = element_blank()
    )

  ggplot2:::print.ggplot(p_density)
  
  # Save plot
  ggsave(
    filename = file.path(charts_dir, "population_density_distribution.png"),
    plot = p_density,
    width = 8,
    height = 4,
    dpi = 300
  )
  
}


```

`r if (in_oxford) "### Age structure"`

`r if (in_oxford) "Using data in Oxford Economics"`

```{r, pop-age-structure, echo = F, message= F, warning=F}
if (in_oxford && exists("pop_dist_structure")) {

  # Print age structure (young, working, elderly) (OUTPUT)
  print_paged_df(pop_dist_structure)

  hues_age_structure <- c("black", "dark grey", "dark grey") %>% #c(scales::hue_pal()(2)) %>% 
    setNames(c("Working", "Young", "65+"))
  hues_age_structure[1] <- "#000000"
  
  pop_dist_plot <- ggplot(pop_dist_structure) +
    geom_area(aes(x = Year, y = Percent, fill = forcats::fct_rev(Group)), alpha = 0.8) +
    geom_text_repel(
      data = subset(pop_dist_structure, Year == 2021),
      aes(label = paste0(Group, "\n", scales::label_percent()(Percent)),
          x = 2021, y = (pct_sum - Percent/2), color = forcats::fct_rev(Group)), 
      hjust = 0, size = 2.5, xlim = c(2021.25, 2025), box.padding = 0.05,
      segment.linetype = "dotted") + 
    scale_y_continuous(labels = scales::label_percent(), expand = c(0,0)) + 
    scale_x_continuous(
      minor_breaks = 2000:2025,
      breaks = c(2000, 2005, 2010, 2015, 2020),
      limits = c(2000, 2025), expand = c(0,0)) +
    scale_fill_manual(values = hues_age_structure) +
    scale_color_manual(values = hues_age_structure) +
    theme_minimal() +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          legend.position = "none") +
    labs(title = "Age Structure, 2000-2021",
         y = "Percent of Population",
         x = "Year")
  ggsave("03-render-output/plots/oxford-age-structure-area.png", plot = pop_dist_plot, device = "png",
         width = 4, height = 3.5, units = "in", dpi = "print")
  
  # Print plot of age structure (OUTPUT)
  ggplot2:::print.ggplot(pop_dist_plot)
}


```

### Share of GDP, Emp, Pop
`r if (!in_oxford) "_Only included if city is in Oxford Economics database. City is not._"`
```{r, national_shares-plot, echo = F, message= F, warning=F}
if (in_oxford && exists("national_shares_long")) {
 
  national_shares_plot <- national_shares_long %>%
    ggplot(aes(x = Percentage, y = Location, fill = Indicator)) +
    geom_col(position = "dodge") +
    # geom_text(aes(label = round(100*Percentage)), position = position_dodge(width = 1))
    geom_text(data = filter(national_shares_long, Group == city), 
              aes(x = Percentage, y = Location, group = Indicator,
                  label = paste0("------", Indicator, ": ", round(Percentage * 100,1),"%")),
              position = position_dodge(width = 1), hjust = 0, size = 2.5) +
    theme_minimal() +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
          panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray"),
          panel.grid.major.y = element_blank(),
          legend.position = "bottom",
          legend.key.height = unit(1/8, "in"),
          axis.title.y = element_blank()) +
    labs(title = "Cities' shares of national population, employment & GDP",
         y = "City",
         fill = "") + 
    scale_x_continuous(breaks = seq(0, 1, .05),
                       labels = scales::percent_format(accuracy = 1),
                       expand = expansion(c(0, NA))) +
    # Keeps text on plot when city's share is one of largest
    expand_limits(x = 1.2 * max(filter(national_shares_long, Group == city)$Percentage))
  # scale_fill_manual(values = c("GDP Share" = "dark grey", "Employment Share" = "light grey", "Population Share" = "black")) +
  # scale_color_manual(values = c("GDP Share" = "dark grey", "Employment Share" = "black", "Population Share" = "black"))
  ggsave("03-render-output/plots/oxford-national-shares.png", plot = national_shares_plot, device = "png",
         width = 8, height = 6.5, units = "in", dpi = "print")
  ggplot2:::print.ggplot(national_shares_plot)
}
```

### GDP, Pop, Emp Growth
`r if (!in_oxford) "_Only included if city is in Oxford Economics database. City is not._"`
```{r growth-tables, echo = F, message= F, warning=F}
if (in_oxford) {
  pop_growth_years <- oxford %>% subset(Indicator == "Total population") %>%
    pivot_longer(cols = contains("20"), values_to = "Value", names_to = "Year") %>%
    mutate(Group, Location, Year = as.numeric(Year), Value, .keep = "used") %>%
    subset(between(Year, 2000, 2021)) %>%
    group_by(Group, Location) %>%
    mutate(Growth = Value / lag(Value) - 1) %>% 
    slice_max(Year, n = 20)
  pop_growth2 <- pop_growth_years %>%
    summarize(`Population Growth` = mean(Growth, na.rm = T), .groups = "drop") %>%
    arrange(desc(`Population Growth`))
  print_text("Population growth from 2000 to 2021")
  print_paged_df(pop_growth2)

  print_text("Employment growth from 2000 to 2021")
  print_paged_df(emp_growth)
  
  print_text("Employment growth in focus city for each year")
  filter(emp_growth_years, Location == city) %>% print_paged_df()


  print_text("GDP growth from 2000 to 2021")
  print_paged_df(gdp_growth)
  # filter(gdp_growth_years, Year == 2022) %>% arrange(desc(Growth))
  print_text("GDP growth in focus city for each year")
  print_paged_df(filter(gdp_growth_years, Location == city))

  # city_pop_growth <- subset(pop_growth2, Location == city, select = `Population Growth`) %>% pull()
  # ggplot(pop_growth2) +
  #   geom_boxplot(aes(y = `Population Growth`), fill = "#00BFC5") +
  #   theme_minimal() +
  #   theme(axis.title.x = element_blank(),
  #         axis.text.x = element_blank(),
  #         axis.ticks.x = element_blank()) +
  #   labs(title = paste("Population Growth of", city, "and benchmark cities"), x = "") +
  #   geom_label(data = data.frame(x = 0, y = city_pop_growth, name = city),
  #              aes(x = x, y = y, label = name), label.size = NA) +
  #   scale_y_continuous(breaks = seq(0, .2, .005),
  #                      minor_breaks = seq(0, .2, .001),
  #                      labels = scales::percent_format(accuracy = 0.1))
  # ggsave("03-render-output/plots/oxford-pop-growth-box.png", device = "png",
  #        width = 8, height = 5, units = "in", dpi = "print")
}


```

```{r pop-growth-line, echo = F, message= F, warning=F}
print_text("Total pop in latest year")
pop_longitude %>% group_by(Location) %>% slice_max(Year) %>% arrange(-Population) %>% select(Group, Location, Year, Population) %>%
  arrange(Group, Population) %>%
  print_paged_df()

multi_city_pop_plot <- ggplot(pop_longitude) +
  geom_line(aes(x = Year, y = `Population`, group = Location, color = Group)) +
  scale_color_manual(values = hues) +
  theme_minimal() +
  theme(panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
        panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray"),
        axis.line = element_line(linewidth = .5, color = "black"),
        legend.position = "none") +
  labs(title = paste("Population of", city, "and benchmark cities"), x = "") +
  scale_y_continuous(limits = c(0, NA), labels = scales::label_comma(scale_cut = scales::cut_short_scale()), expand = c(0,0)) +
  scale_x_continuous(
    breaks = seq(min(pop_longitude$Year), 2021, 5),
    minor_breaks = seq(2000,2030,1), expand = c(0,0)) +
  # geom_dl(aes(label = Location), method = list(dl.combine("first.points", "last.points")), cex = 0.8) 
  geom_text_repel(
    data = \(x) slice_max(x, order_by = Year, by = Location), aes(label = Location, x = 2021, y = `Population`, color = Group), 
    hjust = 0, size = 3, xlim = c(2021.25, 2030), box.padding = 0.05,
    direction = "y", segment.linetype = "dotted") +
  coord_cartesian(xlim = c(max(2000, min(c(pop_longitude$Year))), 2028))

ggplot2:::print.ggplot(multi_city_pop_plot)
ggsave("03-render-output/plots/oxford-pop_line.png", device = "png",
       width = 5, height = 4.6, units = "in", dpi = "print")
# print_paged_df(pop_longitude)


```

```{r employment-lineplot, echo = F, message= F, warning=F}
if (in_oxford && exists("emp_longitude")) {
  
  print_text("Total employed in 2021")
  emp_longitude %>% slice_max(Year) %>% arrange(-`Total employed`) %>% select(Group, Location, `Total employed`) %>%
    print_paged_df()
  
  multi_city_emp_plot <- ggplot(emp_longitude) +
    geom_line(aes(x = Year, y = `Total employed`, group = Location, color = Group)) +
    scale_color_manual(values = hues) +
    theme_minimal() +
    theme(panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
          panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray"),
          axis.line = element_line(linewidth = .5, color = "black"),
          legend.position = "none") +
    labs(title = paste("Employment of", city, "and benchmark cities"), x = "") +
    scale_y_continuous(
      limits = c(0, NA), expand = c(0,0),
      labels = scales::label_comma(scale_cut = scales::cut_short_scale())) +
    scale_x_continuous(breaks = seq(min(emp_longitude$Year), 2021, 5), minor_breaks = seq(2000,2030,1), limits = c(min(emp_longitude$Year), 2028), expand = c(0,0)) +
    # geom_dl(aes(label = Location), method = list(dl.combine("first.points", "last.points")), cex = 0.8) 
    geom_text_repel(
      data = \(x) subset(x, Year == 2021),
      aes(label = Location, x = 2021, y = `Total employed`, color = Group), 
      hjust = 0, size = 3, xlim = c(2021.25, 2030), box.padding = 0.05,
      direction = "y", segment.linetype = "dotted")
  ggplot2:::print.ggplot(multi_city_emp_plot)
  ggsave("03-render-output/plots/oxford-emp_line.png", device = "png", plot = multi_city_emp_plot,
         width = 5, height = 4.6, units = "in", dpi = "print")
}
```

```{r gdp-lineplot, echo = F, message= F, warning=F}
if (in_oxford && exists("gdp_longitude")) {
  
  print_text("Total GDP in 2021")
  gdp_longitude %>% slice_max(Year) %>% arrange(-GDP) %>% select(Group, Location, GDP) %>%
    print_paged_df()
  
  multi_city_gdp_plot <- ggplot(gdp_longitude) +
    geom_line(aes(x = Year, y = `GDP`, group = Location, color = Group)) +
    scale_color_manual(values = hues) +
    theme_minimal() +
    theme(panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
          panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray"),
          axis.line = element_line(linewidth = .5, color = "black"),
          legend.position = "none") +
    labs(title = paste("GDP of", city, "and benchmark cities"), x = "") +
scale_y_continuous(
      limits = c(0, NA), expand = c(0,0),
      labels = scales::label_dollar(scale_cut = scales::cut_short_scale())) +
    scale_x_continuous(
      breaks = seq(min(gdp_longitude$Year), 2021, 5),
      minor_breaks = seq(2000,2030,1),
      limits = c(min(gdp_longitude$Year), 2028), expand = c(0,0)) +
    # geom_dl(aes(label = Location), method = list(dl.combine("first.points", "last.points")), cex = 0.8) 
    geom_text_repel(data = \(x) subset(x, Year == 2021), aes(label = Location, x = 2022, y = `GDP`, color = Group), 
                    hjust = 0, size = 3, xlim = c(2021.25, 2030), box.padding = 0.05, 
                    direction = "y", segment.linetype = "dotted")
  ggplot2:::print.ggplot(multi_city_gdp_plot)
  ggsave("03-render-output/plots/oxford-gdp_line.png", device = "png",
         width = 5, height = 4.6, units = "in", dpi = "print")
}


```

### GDP per capita
`r if (!in_oxford) "_Only included if city is in Oxford Economics database. City is not._"`
```{r gdp-per-capita, echo = F, message= F, warning=F}
if (in_oxford && exists("gdppc")) {

  gdppc_plot <- ggplot(gdppc) +
    geom_col(aes(y = fct_reorder(Location, -`GDP per capita`), x = `GDP per capita`, fill = Group)) +
    theme_minimal() +
    theme(panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
          panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray"),
          axis.line = element_line(linewidth = .5, color = "black"),
          legend.position = "none") +
    labs(title = paste("GDP per capita in", city, "and benchmark cities"),
         y = "City") + 
    scale_x_continuous(labels = scales::dollar_format(scale_cut = scales::cut_short_scale()), expand = expansion(mult = c(0,0.05)))
  ggplot2:::print.ggplot(gdppc_plot)
  ggsave("03-render-output/plots/oxford-gdp-per-capita.png", plot = gdppc_plot, device = "png",
         width = 8, height = 5, units = "in", dpi = "print")

  print_paged_df(subset(gdppc, Location == city))

  gdppc_longitude %>%
      slice_max(Year, n = 10) %>%
      summarize(`growth_pct` = mean(growth_pct), .groups = "drop") %>%
      arrange(desc(`growth_pct`)) %>%
      print_paged_df()

  gdppc_longitude_plot <- ggplot(gdppc_longitude) +
    geom_line(aes(x = Year, y = `GDP per capita`, group = Location, color = Group)) +
    scale_color_manual(values = hues) +
    theme_minimal() +
    theme(panel.grid.major = element_line(linewidth = .125, color = "dark gray"),
          panel.grid.minor = element_line(linewidth = .125, linetype = 2, color = "dark gray"),
          axis.line = element_line(linewidth = .5, color = "black"),
          legend.position = "none") +
    labs(title = paste("GDP per capita of", city, "and benchmark cities"), x = "") +
    scale_y_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale()), expand = c(0,0)) +
    scale_x_continuous(breaks = seq(min(gdppc_longitude$Year), 2021, 5), minor_breaks = seq(2000,2030,1), limits = c(min(gdppc_longitude$Year), 2027), expand = c(0,0)) +
    # geom_dl(aes(label = Location), method = list(dl.combine("first.points", "last.points")), cex = 0.8) 
    geom_text_repel(
      data = \(x) subset(x, Year == 2021), aes(label = Location, x = 2021, y = `GDP per capita`, color = Group), 
      hjust = 0, size = 3, xlim = c(2021.25, 2030), box.padding = 0.05,
      direction = "y", segment.linetype = "dotted")

  ggplot2:::print.ggplot(gdppc_longitude_plot)

  ggsave("03-render-output/plots/oxford-gdp-per-capita_line.png", gdppc_longitude_plot, device = "png",
         width = 8, height = 6.5, units = "in", dpi = "print")


}
```

### Share of employment be sector
`r if (!in_oxford) "_Only included if city is in Oxford Economics database. City is not._"`
```{r share-employment-sectors, echo = F, message= F, warning=F}

if (in_oxford && exists("emp_shares")) {

  print_text("Sectors by employment")
  print_paged_df(emp_shares2)
  
  emp_shares_plot <- emp_shares2 %>%
    ggplot() +
    geom_col(aes(x = Indicator, y = Share, fill = Group),
             position = "dodge") +
    # geom_text_repel(data = emp_shares, aes(x = Indicator, y = Share, label = Location), direction = "y", hjust = 0, nudge_x = 0.1, size = 2) +
    # geom_text(data = emp_shares, aes(x = Indicator, y = Share, label = Location), hjust = 0, nudge_x = 0.1, size = 2) +
    geom_point(data = emp_shares %>% filter(Location != city), aes(x = Indicator, y = Share, group = Group), shape = 21, color = "white", fill="dark grey",  position = position_nudge(x = .25)) +
    # geom_point(data = emp_shares, aes(x = Indicator, y = Share)) +
    scale_y_continuous(
      breaks = seq(0, 1, .05), minor_breaks = seq(0, 1, .01),
      limits = ylims_shares, expand = c(0,0),
      labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual(values = hues) +
    theme_minimal() +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          legend.position = "bottom",
          legend.key.height = unit(1/8, "in")) +
    labs(title = paste("Share of employment by sector in", city, "and median benchmark city"),
         x = "Sector",
         fill = "")
  ggplot2:::print.ggplot(emp_shares_plot)
  ggsave("03-render-output/plots/oxford-employment-sectors.png", plot = emp_shares_plot, device = "png",
         width = 16, height = 5, units = "in", dpi = "print")
         
}
```


```{r share-gva-sectors, echo = F, message= F, warning=F}
if (in_oxford && exists("gva_shares")) {
  print_text("Sectors by GVA")
  print_paged_df(gva_shares2)
  gva_shares_plot <- gva_shares2 %>%
    # arrange(Group, desc(Share)) %>%
    # mutate(Indicator = str_replace(Indicator, "Transport, storage, information & communication services", "Transport & ICT"),
    #        Indicator = factor(Indicator, levels = unique(Indicator))) %>%
    ggplot() +
    geom_col(aes(x = Indicator, y = Share, fill = Group),
             position = "dodge") +
    geom_point(data = gva_shares %>% filter(Location != city), aes(x = Indicator, y = Share, group = Group), shape = 21, color = "white", fill="dark grey", position = position_nudge(x = .25)) +
    scale_y_continuous(breaks = seq(0, 1, .05),
                       minor_breaks = seq(0, 1, .01),
                       limits = ylims_shares,
                       expand = c(0,0),
                       labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual(values = hues) +
    theme_minimal() +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          legend.position = "bottom",
          legend.key.height = unit(1/8, "in")) +
    labs(title = paste("Share of GVA by sector in", city, "and median benchmark city"),
         x = "Sector",
         fill = "")
  ggplot2:::print.ggplot(gva_shares_plot)
  ggsave("03-render-output/plots/oxford-gva-sectors.png", plot = gva_shares_plot, device = "png",
         width = 16, height = 5, units = "in", dpi = "print")
}

```

### Economic Inequality
`r if (!in_oxford) "_Only included if city is in Oxford Economics database. City is not._"`
```{r incomes-plots, echo = F, message= F, warning=F}
if (in_oxford && exists("incomes")) {

  # subset(oxford, !is.na(`Income Band`)) %>% select(`Income Band`) %>% unique() %>% subset(str_detect(`Income Band`, "PPP constant 20"))
  
  household_incomes_bar_plot <- incomes %>% group_by(Location)  %>%
    subset(Year == 2021) %>%
    mutate(Percent = Households / sum(Households), sum = sum(Households)) %>%
    ungroup() %>%
    group_by(Group, Band) %>%
    summarize(Percent = median(Percent), .groups = "drop") %>%
    ggplot(aes(y = Band, x = Percent, fill = Group)) +
    geom_col(position = "dodge") +
    scale_fill_manual(values = hues) +
    scale_x_continuous(labels = scales::label_percent()) + 
    theme_minimal() +
    theme(axis.line = element_line(linewidth = .5, color = "black")) +
    # legend.position = "bottom") +
    labs(title = "Household income distribution (PPP constant 2015 dollars)",
         x = "Percent of Households",
         y = "Income Group")
  ggsave("03-render-output/plots/oxford-household-incomes-bar.png", plot = household_incomes_bar_plot, device = "png",
         width = 8, height = 5, units = "in", dpi = "print")
  ggplot2:::print.ggplot(household_incomes_bar_plot)
  
  
  household_incomes_area_plot <- ggplot(incomes_city2) +
    geom_area(aes(x = Year, y = Percent, fill = forcats::fct_rev(Band)), color = "black", alpha = 0.8) +
    geom_text_repel(
      data = \(x) subset(x, Year == 2021),
      aes(label = Band, x = 2021, y = (pct_sum - Percent/2), color = forcats::fct_rev(Band)), 
      hjust = 0, size = 3, xlim = c(2021.25, 2030), box.padding = 0.02) +
    scale_y_continuous(labels = scales::label_percent(),
                       expand = expansion(mult = c(0,0))) + 
    scale_x_continuous(minor_breaks = 2002:2030, limits = c(min(incomes_city2$Year), max(incomes_city2$Year) + 3),
                       expand = expansion(mult = c(0,0))) + 
    theme_minimal() +
    theme(axis.line = element_line(linewidth = .5, color = "black"), legend.position = "none") +
    labs(title = paste0("Household income distribution, ", min(incomes_city2$Year), "-2021 (PPP constant 2015 dollars)"),
         y = "Percent of Households",
         x = "Year")
  ggplot2:::print.ggplot(household_incomes_area_plot)
  ggsave("03-render-output/plots/oxford-household-incomes-area.png", plot = household_incomes_area_plot, device = "png",
         # width = 16, height = 3.4, units = "in", dpi = "print")
         width = 4, height = 3.5, units = "in", dpi = "print")
}

```

## WSF
### Urban built-up area
```{r urban-built, echo = F, message= F, warning=F}

if (!exists('wsf')) print_text(paste_bold("No WSF file available")) else {
  # Urban built-up area ----
  if (str_detect(wsf_file, "xlsx$")) wsf <- read_xlsx(wsf_file, sheet = "WSF")
  if (str_detect(wsf_file, "csv$")) wsf <- read_csv(wsf_file, col_types = "dd")
  wsf <- wsf %>%
    rename(Year = 1) %>%
    select(Year, uba_km2 = any_of(c("AREA_sq_km", "cumulative sq km"))) %>%
    mutate(growth_pct = (uba_km2 / lag(uba_km2) - 1), growth_km2 = uba_km2 - lag(uba_km2))
  uba_plot <- wsf %>%
    ggplot +
    geom_line(aes(x = Year, y = uba_km2)) +
    scale_x_continuous(
      breaks = seq(1985, 2020, 5),
      minor_breaks = seq(1985, 2021, 1)) + 
    scale_y_continuous(labels = scales::comma, limits = c(0, max(wsf$uba_km2)), expand = c(0, NA)) +
    theme_minimal() +
    labs(title = "",#paste(city, "Urban Built-up Area, 1985-2015"),
         y = bquote('Urban built-up area,'~km^2)) +
    theme(axis.line = element_line(linewidth = .5, color = "black"))
  ggsave("03-render-output/plots/wsf-uba-plot.png", plot = uba_plot, device = "png",
         width = 4, height = 3.5, units = "in", dpi = "print")
  ggplot2:::print.ggplot(uba_plot)
  
  first_area <- wsf$uba_km2[1]
  latest_area <- tail(wsf$uba_km2, 1)
  first_year <- wsf$Year[1]
  latest_year <- tail(wsf$Year, 1)
  pct_growth <- 100 * (latest_area - first_area)/first_area

  print_text(paste0("\nThe city's built-up area grew from ", paste_bold(round(first_area, 2)), " km² in ", first_year, " to ", paste_bold(round(latest_area, 2)), " in ", paste_bold(latest_year), " for ", paste_bold(round(pct_growth, 2)), "% growth"))
  
  print_paged_df(wsf)
  write_csv(wsf, "02-process-output/tabular/urban-extent-and-change.csv")
}

```

### WSF Tracker

```{r wsf-tracker, echo = F, message= F, warning=F}

if (!exists('wsf_tracker')) print_text(paste_bold("No WSF tracker file available")) else {
  # Urban built-up area ----
  if (str_detect(wsf_tracker_file, "xlsx$")) wsf_tracker <- read_xlsx(wsf_tracker_file, sheet = "WSF")
  if (str_detect(wsf_tracker_file, "csv$")) wsf_tracker <- read_csv(wsf_tracker_file, col_types = "cdd")
  wsf_tracker <- wsf_tracker %>%
    # mutate(date = zoo::as.yearmon(date)) %>%
    mutate(date = zoo::as.yearmon(paste(year, month, sep = "-"))) %>%
    select(date, uba_km2 = any_of(c("AREA_sq_km", "cumulative sq km"))) %>%
    mutate(growth_pct = (uba_km2 / lag(uba_km2) - 1), growth_km2 = uba_km2 - lag(uba_km2))
  uba_tracker_plot <- wsf_tracker %>%
    ggplot() +
    geom_line(aes(x = date, y = uba_km2)) +
    scale_x_yearmon(breaks = 2016.5 + 0:4*2) +
    scale_y_continuous(labels = scales::comma, limits = c(0, NA), expand = expansion(c(0, 0.05))) +
    theme_minimal() +
    labs(title = "",#paste(city, "Urban Built-up Area, 1985-2015"),
         y = bquote('Urban built-up area,'~km^2)) +
    theme(
      axis.line = element_line(linewidth = .5, color = "black"),
      axis.title.x = element_blank())
  ggsave("03-render-output/plots/wsf-uba-tracker-plot.png", plot = uba_tracker_plot, device = "png",
         width = 4, height = 3.5, units = "in", dpi = "print")
  ggplot2:::print.ggplot(uba_tracker_plot)
  
  first_area <- wsf_tracker$uba_km2[1]
  latest_area <- tail(wsf_tracker$uba_km2, 1)
  first_date <- wsf_tracker$date[1]
  latest_date <- tail(wsf_tracker$date, 1)
  pct_growth <- 100 * (latest_area - first_area)/first_area
  print_text(paste0("\nThe city's built-up area grew from ", paste_bold(round(first_area, 2)), " km² in ", paste_bold(first_date), " to ", paste_bold(round(latest_area, 2)), " in ", paste_bold(latest_date), " for ", paste_bold(round(pct_growth, 2)), "% growth"))
  
  print_paged_df(wsf_tracker)
  write_csv(wsf_tracker, "02-process-output/tabular/urban-extent-and-change-tracker.csv")
}


```

### Landcover
```{r landcover, echo = F, message= F, warning=F}

if (!exists('lc')) print_text(paste_bold("No Landcover file available")) else {

  print_paged_df(lc)
# lc %>% arrange(desc(Percent)) %>% View()

  lc_colors <- c(
    "Tree cover" = "#397e48",
    "Built-up" = "#c4281b",
    "Grassland" = "#88af52",
    "Bare / sparse vegetation" = "#a59b8f",
    "Cropland" = "#e49634",
    "Permanent water bodies" = "#429bdf",
    "Shrubland" = "#dfc25a",
    "Herbaceous wetland" = "#7d87c4",
    "Mangroves" = "#00cf75")

  lc_plot <- ggdonut(lc, "Land Cover", "Percent", colors = lc_colors, title = "Land Cover")
  ggplot2:::print.ggplot(lc_plot)
  ggsave("03-render-output/plots/wsf-lc-pieplot.png", plot = lc_plot, device = "png",
        width = 8, height = 5, units = "in", dpi = "print")


  
# lc %>%
#   ggplot() +
#   geom_col(aes(x = `Land Cover`, y = Percent, fill = `Land Cover`)) +
#   theme_minimal() +
#   theme(
#     legend.position = "none",
#     axis.line = element_line(linewidth = .5, color = "black")) +
#   scale_y_continuous(breaks = seq(0, 1, .1),
#                      minor_breaks = seq(0, 1, .05),
#                      expand = c(0,0),
#                      labels = scales::percent_format(accuracy = 1)) +
#   scale_fill_manual(values = lc_colors) +
#   labs(title = paste("Land Cover by Percent of Urban Area", city),
#        y = "Percent of Urban Area")
# ggsave("03-render-output/plots/wsf-lc-barplot.png", device = "png",
#        width = 8, height = 5, units = "in", dpi = "print")

# install.packages("treemapify")
# library(treemapify)
# lc %>%
#   ggplot(aes(fill = `Land Cover`, area = Percent, label = `Land Cover`)) +
#   geom_treemap() +
#   geom_treemap_text(color = "white", place = "centre") +
#   scale_fill_manual(values = lc_colors) +
#   theme_minimal() +
#   # theme(axis.line = element_line(linewidth = .5, color = "black")) +
#   # scale_y_continuous(breaks = seq(0, 1, .1),
#   #                    minor_breaks = seq(0, 1, .05),
#   #                    expand = c(0,0),
#   #                    labels = scales::percent_format(accuracy = 1))
#   labs(title = paste("Land Cover by Percent of Urban Area", city)) +
#   theme(legend.position = "none")
# ggsave("03-render-output/plots/oxford-lc-treeplot.png", device = "png",
#        width = 8, height = 8, units = "in", dpi = "print")

# lc %>%
#   ggplot(aes(x = "", fill = `Land Cover`, y = Percent, label = `Land Cover`)) +
#   geom_col(width = 1, color = "white") +
#   scale_fill_manual(values = lc_colors) +
#   # xlim(c(2,4)) +
#   coord_polar("y", start = 0) +
#   theme_void() +
#   # theme(axis.line = element_line(linewidth = .5, color = "black")) +
#   # scale_y_continuous(breaks = seq(0, 1, .1),
#   #                    minor_breaks = seq(0, 1, .05),
#   #                    expand = c(0,0),
#   #                    labels = scales::percent_format(accuracy = 1))
#   labs(title = paste("Land Cover by Percent of Urban Area", city)) +
#   theme(legend.position = "none"

  lc1 <- lc$`Land Cover`[1]
  lc1_pct <- lc$Percent[1] %>% scales::percent(0.1)
  lc2 <- lc$`Land Cover`[2]
  lc2_pct <- lc$Percent[2] %>% scales::percent(0.1)
  lc3 <- lc$`Land Cover`[3]
  lc3_pct <- lc$Percent[3] %>% scales::percent(0.1)

  print("\n")
  print_text(paste0(paste_bold(lc1), " comprises the majority of land in the ", paste_bold(city), " area of interest (", paste_bold(lc1_pct), ")."))
  print_text(paste0("The next largest land classes are ", paste_bold( lc2), " (", paste_bold(lc2_pct), ") and ", paste_bold(lc3), " (", paste_bold(lc3_pct), ")"))

}


```



### Elevation
```{r elevation-donut, echo = F, message= F, warning=F}


if (!exists('elevation')) print_text(paste_bold("No Elevation file available")) else {

  # elevation2 %>%
  #   ggplot() +
  #   geom_step(aes(x = Elevation, y = Percent), color = "#00BFC5", fill = "#00BFC5") +
  #   theme_minimal() +
  #   theme(axis.line = element_line(linewidth = .5, color = "black")) +
  #   scale_y_continuous(breaks = seq(0, 1, .1),
  #                      minor_breaks = seq(0, 1, .05),
  #                      expand = c(0,0),
  #                      labels = scales::percent_format(accuracy = 1)) +
  #   scale_x_continuous(breaks = c(0,320,650,970,1300,1620),
  #                      expand = c(0,0),
  #                      labels = scales::comma) +
  #   labs(title = paste("Elevation of", city),
  #        y = "Percent of Urban Area",
  #        x = "Elevation (m)")

  elevation_names <- elevation$Bin
  elevation_colors <- c(
    "#F2D7D3",
    "#DCBAC5",
    "#C59FB8",
    "#AF84A9",
    "#97679C") %>%
    setNames(elevation_names)
  
  elevation_plot <- ggdonut(elevation, "Bin", "percent", colors = elevation_colors, title = "Elevation")

  ggplot2:::print.ggplot(elevation_plot)
  ggsave("03-render-output/plots/wsf-elevation-pieplot.png", device = "png",
        width = 8, height = 5, units = "in", dpi = "print")

  # elevation2 %>%
  #   ggplot() +
  #   geom_col(aes(x = legend, y = Percent,  fill = legend)) +
  #   theme_minimal() +
  #   theme(
  #     legend.position = "none",
  #     axis.line = element_line(linewidth = .5, color = "black")) +
  #   scale_y_continuous(breaks = seq(0, 1, .1),
  #                      minor_breaks = seq(0, 1, .05),
  #                      expand = c(0,0),
  #                      labels = scales::percent_format(accuracy = 1)) +
  #   # scale_x_continuous(breaks = c(0,320,650,970,1300,1620),
  #   #                    expand = c(0,0),
  #   #                    labels = scales::comma) +
  #   scale_fill_manual(values = elevation_colors) +
  #   labs(title = paste("Elevation of", city),
  #        y = "Percent of Urban Area",
  #        x = "Elevation (m)")
  # ggsave("03-render-output/plots/oxford-elevation-barplot.png", device = "png",
  #        width = 8, height = 5, units = "in", dpi = "print")
  # 
  # elevation_step <- rbind(elevation2, c(NA, 0, 0, 0, 1620))
  # elevation_step <- bind_rows(old = elevation_step, 
  #                             new = elevation_step %>% mutate(Percent = lag(Percent)),
  #                             .id = "source") %>%
  #   arrange(Elevation, source)
  # ggplot(elevation_step, aes(Elevation, Percent)) + 
  #   geom_ribbon(aes(x = Elevation, ymin = 0, ymax = Percent), data = elevation_step) +
  #   theme_minimal() +
  #   theme(axis.line = element_line(linewidth = .5, color = "black")) +
  #   scale_y_continuous(breaks = seq(0, 1, .1),
  #                      minor_breaks = seq(0, 1, .05),
  #                      expand = expansion(mult = c(0, .05)),
  #                      labels = scales::percent_format(accuracy = 1)) +
  #   scale_x_continuous(breaks = c(0,320,650,970,1300,1620),
  #                      expand = expansion(mult = c(0, .05)),
  #                      labels = scales::comma) +
  #   labs(title = paste("Elevation of", city),
  #        y = "Percent of Urban Area",
  #        x = "Elevation (m)")
  # ggsave("03-render-output/plots/oxford-elevation-stepplot.png", device = "png",
  #        width = 8, height = 8, units = "in", dpi = "print")

}
```

### Slope
```{r slope-donut, echo = F, message= F, warning=F}

if (!exists('slope')) print_text(paste_bold("No Slope file available")) else {

  slope_names <- arrange(slope, Slope)$Bin
  slope_colors <- c(
    "#F9F9DF",
    "#E8D6B9",
    "#D5B495",
    "#BE9172",
    "#A86E50") %>%
    setNames(slope_names)

  slope_plot <- ggdonut(slope, "Bin", "percent", colors = slope_colors, title = "Slope")
  ggplot2:::print.ggplot(slope_plot)
  ggsave("03-render-output/plots/wsf-slope-pieplot.png", device = "png",
        width = 8, height = 5, units = "in", dpi = "print")

}

```

## Urban Flooding
### Fluvial


```{r fluvial-flood, echo = F, message= F, warning=F}

if (!exists('fu')) print_text(paste_bold('No Fluvial flood data available')) else {
  
  fu_plot <- fu %>%
    ggplot() +
    geom_line(aes(x = Year, y = uba_km2_exposed)) +
    scale_x_continuous(
      breaks = seq(1985, 2020, 5),
      minor_breaks = seq(1985, 2021, 1)) + 
    scale_y_continuous(labels = scales::comma, limits = c(0, NA), expand = expansion(c(0, 0.05))) +
    theme_minimal() +
    labs(title = "",#paste(city, "Built-Up Area Exposed to River Flooding Historical Growth, 1985-2015"),
         y = bquote('Exposed'~km^2)) +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          axis.title.x = element_blank())
          
  ggplot2:::print.ggplot(fu_plot)

  ggsave("03-render-output/plots/wsf-fu-plot.png", plot = fu_plot, device = "png",
         width = 4, height = 3.5, units = "in", dpi = "print")

  print_paged_df(bind_rows(head(fu, 1), tail(fu, 1)))

  print_text(glue("In {paste_bold(fu$Year[1])}, {paste_bold(round(fu$uba_km2_exposed[1], 2)) } km² of the built-up area ({paste_bold(fu$percent_uba_exposed[1])}) was exposed to river flooding at a minimum depth of 15 cm"))
  print_text(glue("In { paste_bold(tail(fu$Year, 1))}, {paste_bold(round(tail(fu$uba_km2_exposed, 1), 2))} km² of the built-up area ({paste_bold(tail(fu$percent_uba_exposed, 1))}) was exposed"))

print_text(flood_pop_area("fluvial"))
print_text(flood_string("fluvial"))

}



```

### Pluvial
```{r pluvial-flood, echo = F, message= F, warning=F}

if (!exists('pu')) print_text(paste_bold('No Pluvial flood data available')) else {
  
  pu_plot <- pu %>%
    ggplot +
    geom_line(aes(x = Year, y = uba_km2_exposed)) +
    scale_x_continuous(
      breaks = seq(1985, 2020, 5),
      minor_breaks = seq(1985, 2021, 1)) + 
    scale_y_continuous(labels = scales::comma, limits = c(0, NA), expand = expansion(c(0, 0.05))) +
    theme_minimal() +
    labs(title = "",#paste(city, "Built-Up Area Exposed to Rainwater Flooding Historical Growth, 1985-2015"),
         y = bquote('Exposed'~km^2)) +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          axis.title.x = element_blank())
  ggplot2:::print.ggplot(pu_plot)
  ggsave("03-render-output/plots/wsf-pu-plot.png", plot = pu_plot, device = "png",
         width = 4, height = 3.5, units = "in", dpi = "print")
  
  print_paged_df(bind_rows(head(pu, 1), tail(pu, 1)))

  print_text(glue("In {paste_bold(pu$Year[1])}, {paste_bold(round(pu$uba_km2_exposed[1], 2))} km² of the built-up area ({paste_bold(pu$percent_uba_exposed[1])}) was exposed to river flooding at a minimum depth of 15 cm"))
  print_text(glue("In {paste_bold(tail(pu$Year, 1))}, {paste_bold(round(tail(pu$uba_km2_exposed, 1), 2))} km² of the built-up area ({paste_bold(tail(pu$percent_uba_exposed, 1))}) was exposed"))
  # pu %>% mutate(diff = pu$uba_km2 - lag(pu$uba_km2)) %>% .$diff %>% mean(na.rm = T) 

  print_text(flood_pop_area("pluvial"))
  print_text(flood_string("pluvial"))
}

```

### Coastal
```{r coastal-flood, echo = F, message= F, warning=F}
if (!exists('cu')) print_text(paste_bold('No Coastal flood data available')) else {

  cu_plot <- cu %>%
    ggplot() +
    geom_line(aes(x = Year, y = uba_km2_exposed)) +
    scale_x_continuous(
      breaks = seq(1985, 2020, 5),
      minor_breaks = seq(1985, 2021, 1)) + 
    scale_y_continuous(labels = scales::comma, limits = c(0, NA), expand = expansion(c(0, 0.05))) +
    theme_minimal() +
    labs(title = "",#paste(city, "Built-Up Area Exposed to Coastal Flooding Historical Growth, 1985-2015"),
         y = bquote('Exposed'~km^2)) +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          axis.title.x = element_blank())
  ggplot2:::print.ggplot(cu_plot)
  ggsave("03-render-output/plots/wsf-cu-plot.png", plot = cu_plot, device = "png",
         width = 4, height = 3.5, units = "in", dpi = "print")
  
  print_paged_df(bind_rows(head(cu, 1), tail(cu, 1)))
  print_text(glue("In {paste_bold(cu$Year[1])}, {paste_bold(round(cu$uba_km2_exposed[1], 2))} km² of the built-up area ({paste_bold(cu$percent_uba_exposed[1])}) was exposed to river flooding at a minimum depth of 15 cm"))
  print_text(glue("In {paste_bold(tail(cu$Year, 1))}, {paste_bold(round(tail(cu$uba_km2_exposed, 1), 2))} km² of the built-up area ({paste_bold(tail(cu$percent_uba_exposed, 1))}) was exposed"))

  print_text(flood_pop_area("coastal"))
  print_text(flood_string("coastal"))
}

```

### Combined Fluvial, Pluvial & Coastal
```{r combined-flood, echo = F, message= F, warning=F}

if (!exists('pufu') && !exists('comb') ) print_text(paste_bold('No Combined flood data available')) else {

  pufu_plot <- pufu %>%
    ggplot(aes(x = Year, y = uba_km2_exposed, color = type)) +
    geom_line(data = \(d) filter(d, type == "Combined")) +
    geom_point(data = \(d) filter(d, type != "Combined")) +
    scale_x_continuous(
      breaks = seq(1985, 2020, 5),
      minor_breaks = seq(1985, 2021, 1)) + 
    scale_y_continuous(labels = scales::comma, limits = c(0, NA), expand = expansion(c(0, 0.05))) +
    scale_color_manual(values = c(River = "#F8766D", Rainwater = "#619CFF", Coastal = "#00BA38", Combined = "black")) +
    scale_linetype_manual(values = c(River = "dashed", Rainwater = "dashed", Coastal = "dashed", Combined = "solid")) +
    scale_size_manual(values = c(River = 1, Rainwater = 1, Coastal = 1, Combined = 0.5)) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(title = "",#paste(city, "Built-Up Area Exposed to River & Rainwater Flooding, 1985-2015"),
         y = bquote('Exposed'~km^2), color = "", linetype = "", size = "") +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          axis.title.x = element_blank())
  ggplot2:::print.ggplot(pufu_plot)
  ggsave("03-render-output/plots/wsf-pufu-plot.png", plot = pufu_plot, device = "png",
         width = 4, height = 3.5, units = "in", dpi = "print")
  
  print_paged_df(bind_rows(head(comb, 1), tail(comb, 1)))
  print_text(glue("In {paste_bold(comb$Year[1])}, {paste_bold(round(comb$uba_km2_exposed[1], 2))} km² of the built-up area ({paste_bold(comb$percent_uba_exposed[1])}) was exposed to riverine, surface water, or coastal flooding at a minimum depth of 15 cm"))
  print_text(glue("In {paste_bold(tail(comb$Year, 1))}, {paste_bold(round(tail(comb$uba_km2_exposed, 1), 2))} km² of the built-up area ({paste_bold(tail(comb$percent_uba_exposed, 1))}) was exposed")) 

  print_text(flood_pop_area("comb"))
  print_text(flood_string("comb"))
}



```

## Flood Events
```{r flood-events, echo = F, message= F, warning=F, fig.width=12}
# Download flood archive files (can comment this out to use local file)

if (!exists('flood_archive') || nrow(flood_archive) == 0) print_text(paste_bold('No flood archive data available')) else {
  print_text("Map of floods")
  ggplot(flood_archive) +
    geom_sf(aes(fill = DEAD)) +
    geom_sf(data = aoi)


  print_text("Tally of flood events")

  summarize(floods, count = n(), across(.cols = c(DEAD, DISPLACED), sum), .groups = "drop") %>% 
    print_paged_df()

  print_text("Median flood event")

  floods %>% mutate(
    duration = as.numeric(ENDED - BEGAN)) %>%
    summarize(across(.cols = c(DEAD, DISPLACED, duration), median), .groups = "drop") %>% 
    print_paged_df()

  print_text("Causes of flood events")

  count(floods, MAINCAUSE) %>% print_paged_df()


  month_labels <- cumsum(c("Jan" = 31, "Feb" = 28, "Mar"= 31, "Apr" = 30, "May" = 31, "Jun" = 30,
    "Jul" = 31, "Aug" = 31, "Sep"= 30, "Oct" = 31, "Nov" = 30, "Dec" = 31))

  flood_event_plot <- flood_text_all %>%
    ggplot(aes(y = begin_year, x = begin_day)) +
    geom_text_repel(
      aes(label = text),
      hjust = 0, size = 4, lineheight = 0.9,
      min.segment.length = 0, force_pull = .25, box.padding = .5,
      xlim = c(370, 700), direction = "y",
      segment.color = "grey46", segment.size = .25,
      segment.curvature = 1e-20, segment.square = T) +
    geom_point(aes(size = DISPLACED, alpha = severity), color = "darkblue") +
    scale_x_continuous(
      limits = c(0, 366), expand = c(0, 0),
      breaks = month_labels - 15,
      labels = names(month_labels),
      minor_breaks = c(0, month_labels)) +
    scale_y_reverse(
      limits = c(lubridate::year(Sys.Date()), 1985),
      breaks = seq(1985, 2030, by = 5), minor_breaks = 1985:2030 - 0.5) +
    scale_alpha_discrete(range = c(0.3, 1)) +
    scale_size_continuous(
      range = c(3, 10),
      breaks = range(flood_text_all$DISPLACED),
      labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
    labs(y = "Year", alpha = "Severity", size = "Displaced") +
    theme_minimal() +
    coord_cartesian(clip = 'off') +
    theme(
      axis.line = element_line(linewidth = .5, color = "black"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_line(linewidth = .5, color = "grey92"),
      axis.text = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.position = "bottom",
      plot.margin = margin(5.5, 300, 5.5, 5.5, "pt"),
      plot.background = element_rect(fill = "white", color = NULL))

  ggplot2:::print.ggplot(flood_event_plot)

  ggsave("03-render-output/plots/darthmouth_flood_events.pdf", device = "pdf",
    width = 15, height = 4.36, units = "in", dpi = "print")

}

```

## Earthquakes
```{r earthquake-events, echo = F, message= F, warning=F, fig.width=12}

if (!exists('eq') || nrow(eq) == 0) print_text(paste_bold('No earthquake data available')) else {

  damages <- c("0" = "None",
              "1" = "Limited",
              "2" = "Moderate",
              "3" = "Severe",
              "4" = "Extreme")

  y_ratio <- 51498 / 13514
  # 3.324

  # aoi already loaded in setup.R
  city_point <- aoi %>% st_combine() %>% st_centroid()

  eq_near <- eq %>%
    subset(!is.na(latitude)) %>%
    st_as_sf(coords = c("longitude", "latitude"), crs = st_crs(city_point)) %>%
    mutate(distance = st_distance(geometry, city_point) %>% units::set_units(km),
          magXdist = eqMagnitude * (2000 - units::drop_units(distance))) %>%
    filter(
      (damageAmountOrder >= 2 | damageMillionsDollars >= 1 |
        deaths >= 10 | deathsAmountOrder >= 2 |
        eqMagnitude >= 7.5 |
        intensity >= 10 |
        !is.na(tsunamiEventId)) &
        distance < units::set_units(500, km))

  if (nrow(eq_near) == 0) {
    print_text(paste_bold('No significant earthquakes found within 500 km'))
  } else {

  eq_text <- eq_near %>%
    mutate(
      location = gsub("([\\s-][A-Z])([A-Z]*)", "\\1\\L\\2", locationName, perl = T) %>% 
        str_extract("[^:][\\s]+.*") %>% trimws(),
      fatalities = paste(scales::label_comma()(deaths), "fatalities"),
      fatalities = str_replace_all(
        fatalities, c("1 fatalities" = "1 fatality", "NA fatalities" = "")),
      damage = paste(damages[as.character(damageAmountOrder)], "damage"),
      day = replace_na(day, 1),
      BEGAN = as.Date(paste(year, month, day, sep='-'))) %>%
    arrange(BEGAN) %>%
    mutate(
      line1 = toupper(paste(lubridate::month(BEGAN, label = T, abbr = F), lubridate::year(BEGAN))),
      # line2 = paste0(ifelse(!is.na(location), paste0(location, " "), "") , "(", eqMagnitude, ")"),
      line2 = paste0("M", eqMagnitude, "; ", scales::label_comma(accuracy= 1)(units::drop_units(distance)), " km away"),
      line3 = damage,
      line4 = fatalities,
      text = paste(line1, line2, line3, line4, sep = "; ") %>% str_replace_all(c("; ;" = ";", "; $" = "")),
      above_line = (2*(row_number() %% 2) - 1)*-1) %>%
    select(BEGAN, text, line1, line2, line3, line4, above_line, distance, eqMagnitude, magXdist, location) %>%
    mutate(node_x = BEGAN + y_ratio * 1460*above_line,
          node_y = above_line * (1460 * tanpi(1/6) + 466/2) * y_ratio )

  print_paged_df(select(st_drop_geometry(eq_text), text, location))

  eq_plot <-
    ggplot(eq_text, aes(y = BEGAN, x = drop_units(distance))) +
    # geom_text(data = filter(eq_text, !is.na(eqMagnitude)), aes(size = eqMagnitude, label = eqMagnitude)) +
    geom_text_repel(
      data = \(x) slice_max(x, magXdist, n = 10), aes(label = text),
      size = 4, hjust = 0, direction = "y",
      xlim = c(515, 1000),
      segment.color = "grey46", segment.size = .25,
      segment.curvature = 1e-20, segment.square = T) +
    geom_point(data = \(x) filter(x, !is.na(eqMagnitude)), aes(size = eqMagnitude), shape = 1, color = "black") +
    geom_point(data = \(x) slice_max(x, magXdist, n = 10), aes(size = eqMagnitude), color = "grey46") +
    geom_point(data = \(x) filter(x, is.na(eqMagnitude)), shape = 1, color = "grey92") +
    scale_x_continuous(
      limits = c(0, 500),
      labels = \(x) paste(x, "km"),
      breaks = seq(0, 500, 100),
      expand = expansion(c(0, 0))) +
    scale_y_continuous(
      limits = c(Sys.Date(), as.Date("1900-01-01")),
      # expand = expansion(c(0, 0.05)),
      breaks = seq.Date(as.Date("1900-01-01"), Sys.Date(), by = "20 years"),
      minor_breaks = seq.Date(as.Date("1900-01-01"), as.Date("2030-01-01"), by = "5 years"),
      labels = seq(1900, 2025, by = 20),
      trans = c("date", "reverse")) +
    scale_size(range = c(1, 6)) + # max_size = 7,
    # scale_color_manual(values = c(" " = "grey"), guide = "none") +
    labs(
      x = paste("Distance from", city),
      y = "Year",
      size = "Magnitude",
      # color = "Magnitude unavailable"
      ) +
    theme_minimal() +
    coord_cartesian(clip = 'off') +
    theme(
      axis.line = element_line(linewidth = .5, color = "black"),
      axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.position = "right",
      legend.justification.right = "bottom",
      legend.direction = "horizontal",
      legend.box.margin = margin(0, 0, -37, 15),
      plot.margin = margin(5.5, 300, 5.5, 5.5, "pt"),
      plot.background = element_rect(fill = "white", color = NULL))

  ggsave("03-render-output/plots/ngdc-earthquake-scatter.png", plot = eq_plot, device = "png",
          width = 15, height = 4.36, units = "in", dpi = "print")

  tryCatch(ggplot2:::print.ggplot(eq_plot), error = \(e) warning(paste("Window may be too small for earthquake plot.", e)))
  }
}


```

## Cyclones  
```{r cyclones_near-plot, echo = F, message= F, warning=F}

if (!exists('cyclones_near') || nrow(cyclones_near) == 0) {
  print_text(paste_bold('No cyclone data available'))
} else {
    as.Date("2020-12-31") - as.Date("1984-01-01") # 13514
    as.Date("2025-12-31") - as.Date("1975-01-01") # 18627
    y_ratio <-1.49 # ratio should be 1.378 but doesn't quite work
    
    cyclones_text <- cyclones_near %>%
      st_drop_geometry() %>%
      # group_by(SID, NAME) %>%
      summarize(.by = c(SID, NAME),
        year = min(year),
        month_start = min(month),
        month_start_name = month.name[month_start],
        month = paste(unique(c(min(month), max(month))), collapse = "-"),
        BEGAN = as.Date(paste(year, month_start, "15", sep = "-")),
        USA_SSHS = max(USA_SSHS, na.rm = T),
        min_distance = min(distance_km, na.rm = T)) %>%
      arrange(-USA_SSHS) %>%
      mutate(
        max_scale = case_when(
          USA_SSHS < 0 ~ "Tropical depression",
          USA_SSHS == 0 ~ "Tropical storm",
          T ~ paste("Category", USA_SSHS)),
        text = paste0(toupper(month_start_name), " ", year, "\n", NAME, " - ", max_scale))
        
    cyclones_text_select <- cyclones_text %>% slice_min(min_distance, n = 10) %>%
      arrange(BEGAN) %>%
      mutate(above_line = (2*(row_number() %% 2) - 1)*-1,
        node_x = BEGAN + y_ratio * 1460*above_line,
        node_y = above_line * (1460 * tanpi(1/6) + 466/2) * y_ratio )    
    
    cyclones_lines <- cyclones_text_select %>%
      select(SID, BEGAN, above_line, node_x, node_y) %>%
      mutate(id = row_number()) %>%
      { bind_rows(.,
                  mutate(., node_x = BEGAN, node_y = above_line * 466 / 2),
                  mutate(., node_x = BEGAN, node_y = 0)) }
    
    x_axis_ticks <- tibble(date = seq(as.Date("1980-01-01"),as.Date("2020-12-31"), by = "10 year"), year = lubridate::year(date)) %>%
      bind_rows(tibble(date = seq(as.Date("1985-01-01"),as.Date("2020-12-31"), by = "10 year"), year = NA))
      
    x_axis_years <- data.frame(x = seq(as.Date("1975-01-01"), as.Date("2025-12-31"), by = "years"), y = 0)
    x_axis_months <- data.frame(x = seq(as.Date("1975-01-01"), as.Date("2025-12-31"), by = "2 months"), y = 0)
    
    cyclone_plot <- ggplot(cyclones_text) +
      geom_point(aes(x = BEGAN), y = 0) +
      # geom_point(aes(x = node_x, y = node_y)) +
      geom_text(
        data = cyclones_text_select,
        aes(x = node_x + 0 * y_ratio, y = node_y - 0 * y_ratio, label = text, hjust = (above_line + 1)/2),
        angle = 30, lineheight = 1) +
      scale_x_date(limits = c(as.Date("1974-01-01"),as.Date("2025-12-31")), expand = c(0,0)) +
      scale_y_continuous(limits = c(-1800, 1800) * y_ratio, breaks = c(1825,3650)) +
      geom_path(data = cyclones_lines, aes(x = node_x, y = node_y, group = SID))  +
      geom_point(data = x_axis_years, aes(x = x, y = y), pch = "|", size = 2) +
      geom_point(data = x_axis_months, aes(x = x, y = y), size = 0.25) +
      # geom_hline(yintercept = 0, linetype = "dotted") + 
      geom_point(data = x_axis_ticks, aes(x = date-15, y = 0), pch = "|", size = 4) +
      geom_point(data = x_axis_ticks, aes(x = date+15, y = 0), pch = "|", size = 4) +
      geom_text(data = x_axis_ticks, aes(x = date, y = -0.25 * 466 * y_ratio, label = year)) +
      theme_minimal() +
      theme(axis.text = element_blank(), axis.title = element_blank(), panel.grid = element_blank()) +
      force_panelsizes(rows = unit(5.833, "in"), cols = unit(20, "in"))
    ggsave("03-render-output/plots/ncei-cyclones.pdf", device = "pdf",
          width = 20, height = 5.833, units = "in", dpi = "print")
    
    cyclones_text_select %>% print_paged_df()

    ggplot2:::print.ggplot(cyclone_plot)

    
    cyclones_text %>% filter(min_distance <250) %>% count(USA_SSHS)  %>% print_paged_df()
    cyclones_text %>% mutate(distance_bin = cut_width(min_distance, width = 50, boundary = 0)) %>% count(distance_bin)
}
```



## Photovoltaic Potential (PV)
```{r}

if (!exists('monthly_pv')) print_text(paste_bold('No PV data available')) else {
  monthly_pv %>% print_paged_df()

  pv_plot <- monthly_pv %>%
    ggplot(aes(x = month, y = mean)) +
    annotate("text", x = 1, y = 4.6, label = "Excellent Conditions", vjust = 0, hjust = 0, color = "dark grey") +
    annotate("text", x = 1, y = 3.6, label = "Favorable Conditions", vjust = 0, hjust = 0, color = "dark grey") +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = 1:12, labels = lubridate::month(1:12, label = T) %>% as.character) +
    scale_y_continuous(labels = scales::label_comma(), limits = c(0, NA), expand = expansion(mult = c(0,.05))) +
    labs(title = "Seasonal availability of solar energy",
        x = "Month", y = "Daily PV energy yield (kWh/kWp") +
    geom_hline(yintercept = c(3.5, 4.5), linetype = "dotted") +
    theme_minimal() +
    theme(
      axis.title.x = element_blank(), 
      axis.line = element_line(linewidth = .5, color = "black"),
      panel.grid.minor.x = element_blank())

  ggplot2:::print.ggplot(pv_plot)

  ggsave("03-render-output/plots/monthly-pv.png", device = "png",
        width = 4, height = 3.5, units = "in", dpi = "print")

}

```

## FWI
```{r FWI-plot, echo = F, message= F, warning=F}

month_starts <- cumsum(c("Jan" = 31, "Feb" = 28, "Mar"= 31, "Apr" = 30, "May" = 31, "Jun" = 30,
  "Jul" = 31, "Aug" = 31, "Sep"= 30, "Oct" = 31, "Nov" = 30, "Dec" = 31)/7) - 31/7

if (!exists('fwi')) print_text(paste_bold('No FWI data available')) else {
  
  fwi_plot <- ggplot(fwi, aes(x = week - 1, y = pctile_95)) +
    geom_line() +
    scale_x_continuous(breaks = month_starts + 31/7/2, labels = names(month_starts), minor_breaks = month_starts,
                      expand = c(0,0)) +
    scale_y_continuous(limits = c(0, NA), expand = expansion(c(0, .1))) +   
    #   scale_color_manual(values = hues) +
    theme_minimal() +
    labs(title = paste("FWI in", city, "2016-2021"),
        y = "95th percentile FWI") +
    theme(axis.line = element_line(linewidth = .5, color = "black"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_line(linewidth = .125, color = "dark gray"),
          axis.title.x = element_blank()) +
    geom_vline(xintercept = 0)
  ggplot2:::print.ggplot(fwi_plot)
  ggsave(file.path("03-render-output/plots/nasa-fwi.png"), plot = fwi_plot, device = "png",
        width = 4, height = 3.5, units = "in", dpi = "print")
}

```

## Output Check
```{r output-check, echo = F, message= F, warning=F}
slide_list <- read_csv("source/slide-list.csv", col_types = "cccddcc")

# Check maps
maps <- list.files("03-render-output/maps", pattern = "\\.png$")
missing_maps <- filter(slide_list, filename %ni% maps & type == "map") %>%
  filter(conditions != "none") %>%
  filter(!str_starts(filename, "map_"))

# Check plots (both PNG and PDF)
plots <- list.files("03-render-output/plots", pattern = "\\.(png|pdf)$")
missing_plots <- filter(slide_list, filename %ni% plots & type == "plot") %>%
  filter(conditions != "none")

# Print results
if (nrow(missing_maps) > 0) {
  print_text(paste_bold("Missing Maps:"))
  print_paged_df(missing_maps %>% select(slide, filename, warning_text, data_sources))
}

if (nrow(missing_plots) > 0) {
  print_text(paste_bold("Missing Plots:"))
  print_paged_df(missing_plots %>% select(slide, filename, warning_text, data_sources))
}

if (nrow(missing_maps) == 0 && nrow(missing_plots) == 0) {
  print_text("All expected maps and plots are present.")
}
```