---
title: "CRP—LGCRRP Tutorial"
format: html
editor: 
  markdown: 
    wrap: 72
---

# CRP---LGCRRP Tutorial

Welcome to the CRP---LGCRRP visualizations tutorial!

As you have already seen, creating all of the maps and charts only
requires running `01-main.py` to process the data and `02-main.R` to
visualize it. Here, though, we will briefly walk through what's
happening under the hood so that you can understand what's going on and
know where to make any changes.

1.  Make a plot with `ggplot2`
2.  Make a map with `ggplot2`
3.  Why you don't need to know `ggplot2`
4.  Making edits

## Setup

First, let's load all of the necessary packages using the file
`R/setup.R`. This file is the first thing run by `02-main.R`. It loads
the necessary packages and predefined functions, sets relevant directory
paths, and defines variables that are used throughout the process.

```{r}
source("R/setup.R")
```

## Charts

To make the plots, `02-main.R` calls `R/plots.R`, and this file in turn
calls another file for each chart we are making. Each of these files
make a plot with an R package called `ggplot2`. (The maps are also made
with `ggplot2`, when `02-main.R` calls `R/maps.R`.)

Let's begin with a simple plot of the urban built-up area over time.

The data for this is already processed, and is located in the Cumilla's
`02-process-output/tabular` folder.

```{r}
builtup <- read_csv(file = "mnt/city-directories/cumilla/02-process-output/tabular/cumilla_wsf_stats.csv")
builtup
```

(In R, we define objects using a left arrow: `<-`.)

This CSV contains two columns: the year, from 1985 to 2015, and the
cumulative built-up area of the city.

All `ggplot2` plots begin with a call of the `ggplot()` function. (Note
that there is no "2" in the function name.) Typically, we define the
data inside this function call. We then add (with literal `+`) geometry
functions, which define the variables we want to plot. In this case, `x`
is `year` and y is `cumulative sq km`.

```{r}
ggplot(data = builtup) +
  geom_point(aes(x = year, y = `cumulative sq km`))
  # Note the ` ` around `cumulative sq km`. This is because there are spaces in
  # the variable name, so we need to tell ggplot all of these words go together
```

Above, the geometry function is `geom_point()`; we could alternatively
make a line chart by using `geom_line()`. Other options include
`geom_step()` and `geom_col()`, but there are many more.

```{r}
p <- ggplot(data = builtup) +
  geom_line(aes(x = year, y = `cumulative sq km`))
p
```

We can then start to modify this further, such as changing the labels,
background color, and where the axes begin:

```{r}
builtup_plot <- p + # Note how we defined p in the previous chunk, and are now adding to it
  labs( # Change the labels with labs()
    title = "Urban expansion of Cumilla, 1985-2015",
    x = "Year",
    y = "Urban extent, square km.") +
  scale_y_continuous(
    limits = c(0, NA), # Sets y's lower limit but does not constrain y's upper limit
    expand = expansion(c(0, 0.05)) # How much to expand the axis beyond the limits
    # Also, in R, c() creates a vector of its contents
  ) +
  theme( # Use theme to change aesthetics unrelated to the data
    axis.line = element_line(linewidth = .5, color = "black"),
    panel.background = element_rect(color = NA, fill = "white"),
    panel.grid.major = element_line(color = "lightgrey", linewidth = 0.2)
  )
builtup_plot
```

We can then save this plot using `ggsave()`, which gives us options for
filetype (say, PNG or PDF), and plot dimensions. To make a 16 cm x 10 cm
PNG, we'd do the following.

```{r}
ggsave(
  filename = "mnt/city-directories/cumilla/03-render-output/charts/urban-expansion-tutorial.png",
  plot = builtup_plot,
  width = 15,
  height = 10,
  units = "cm",
  dpi = 300 # This sets the resolution in "dots per inch"
)
```

We can slightly simplify this by using `charts_dir` (a string predefined
in `R/setup.R`) and `file.path()` (a function which combines strings
into a file path).

```{r}
charts_dir
file.path(charts_dir, "urban-expansion-tutorial.png")
```

```{r}
ggsave(
  filename = file.path(charts_dir, "urban-expansion-tutorial.png"),
  plot = builtup_plot,
  width = 15,
  height = 10,
  units = "cm",
  dpi = 300 # This sets the resolution in "dots per inch"
)
```

This is, more or less, the process by which all of the plots are made.
The file `02-main.R` runs a file named `R/plots.R`, which in turn runs a
series of R files which each create a data-specific plot or two, using
similar steps to the above. Some of the files are more complicated, such
as the CCKP plot scripts, which read data from Google Cloud, but this is
the underlying logic. You can modify the plots by changing these files.

#### Takeaways

About the repository - You can modify plots by changing the file that
writes them (e.g., `R/urban-extent.R`) - Doing so will change the plots
for all future versions of that plot - `R/setup.R` has already defined
the path to the charts directory, `charts_dir`, as well as to other
repository directories

About R & ggplot2 - To change plot dimensions, change the `width` and
`height` arguments in `ggsave()` in the relevant file - To change
labels, edit the contents of the `labs()` function - `<-` defines
objects in R - In `ggplot2`, plots begin with `ggplot()` and then
additional items are added with `+` - Data components are added with
geoms, like `geom_point()` and `geom_line()` - Axes and color scales are
modified with scales, like `scale_y_continuous()` or
`scale_x_discrete()`, or `scale_color_manual()` - Use `theme()` to
change non-data aesthetics

About CCKP maps - For the CCKP data, we make two plots for each
variable. One shows the median projections for 3 scenarios (SSP1-2.6,
SSP2-4.5 and SSP3-7.0), while the other shows the median, 10th and 90th
percentiles for SSP3-7.0. To change which SSP is used for the individual
plot, change `selected_ssp` in `R/setup.R` on line 82.

## Maps

### Spatial ggplot2

Maps are also made in `ggplot2`, but there are a few more specific steps
to account for the spatial complexity.

Let's begin, though, by reading in the data, using the terra package's
`rast()` function, which reads raster files. For vector data, we'd use
`vect()`.

```{r}
rast(file.path(spatial_dir, "cumilla_population.tif"))
```

Just as `charts_dir` holds the path to the charts directory (and
`styled_maps_dir` holds the path to the output maps directory),
`spatial_dir` contains the path to where all of the spatial files
reside.

We can further simplify this, by using the pre-defined function
`fuzzy_read()`, which let's us use just part of the file name (so we can
exclude the city name, which, obviously, changes for each city).
`fuzzy_read()` also determines whether the file is a raster file
(GeoTIFF only), or a vector file.

This population file is already in the coordinate reference system EPSG
4326, but to ensure it is, we still project it into it. (Note that the
pipe `%>%` passes whats on the left into the function on the right as
its first argument.)

```{r}
pop <- fuzzy_read(dir = spatial_dir, "population") %>% project(y = "epsg:4326")

pop
```

For raster files, we plot with `ggplot` using `geom_spatraster()`; for
vectors, we use `geom_spatvector()`.

```{r}
ggplot() +
  geom_spatraster(data = pop)
  # Note that for maps, data is defined in geom_spatraster() or geom_spatvector()
  # rather than in ggplot()
```

Here we have our first map!

Now let's change the colors, and stop showing the NA data. We can also
change the legend label with `labs()` and add in the AOI boundary.

```{r}
aoi <- vect(file.path(user_input_dir, "AOI"))
ggplot() +
  geom_spatraster(data = pop) +
  scale_fill_stepsn(
    colors = c("#ECEB72", "#D6BE62", "#BF9152", "#A76642", "#8E3933"),
    na.value = "transparent"
  ) +
  labs(fill = "Population density") +
  geom_spatvector(data = aoi, color = "black", fill = NA)
```

To put everything in context, we add in a base map of raster tiles. The
first time `annotation_map_tile()` is run for a city and zoom level, it
downloads the tiles from the internet and make take a few moments.

While the data is in CRS 4326, the tiles are in CRS 3867, or Web
Mercator. To use the tiles we need to change the projection of the map.
Because the data is in 4326, it is typically safer to define this
projection *after* the data has been included (even though the tiles are
added earlier).

```{r}
tiles <- annotation_map_tile(type = "cartolight", zoom = 15, progress = "none")
ggplot() +
  tiles +
  geom_spatraster(data = pop) +
  scale_fill_stepsn(
    colors = c("#ECEB72", "#D6BE62", "#BF9152", "#A76642", "#8E3933"),
    na.value = "transparent"
  ) +
  labs(fill = "Population density") +
  geom_spatvector(data = aoi, color = "black", fill = NA) +
  coord_sf(crs = "epsg:3857")
```

To finish, let's add a scale bar, north arrow, and adjust some of the
theme styling:

```{r}
tiles <- annotation_map_tile(type = "cartolight", zoom = 15, progress = "none")
ggplot() +
  tiles +
  geom_spatraster(data = pop) +
  scale_fill_stepsn(
    colors = c("#ECEB72", "#D6BE62", "#BF9152", "#A76642", "#8E3933"),
    na.value = "transparent"
  ) +
  labs(fill = "Population density") +
  geom_spatvector(data = aoi, color = "black", fill = NA) +
  coord_sf(crs = "epsg:3857") +
  theme(
    legend.justification = c("left", "bottom"),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  annotation_north_arrow(style = north_arrow_minimal, location = "br", height = unit(1, "cm")) +
  annotation_scale(style = "ticks", aes(unit_category = "metric", width_hint = 0.33), height = unit(0.25, "cm"))
```

#### Takeaways

1.  Just like we use `geom_point()` and `geom_line()` for the charts, we
    use `geom_spatraster()` and `geom_spatvector()` for spatial data
2.  For `geom_spatraster()` and `geom_spatvector()`, `data` must be
    defined within them and not in `ggplot()`
3.  Add basemap tiles with
    `annotation_map_tile(type = "cartolight", zoom = 15, progress = "none")`;
    these need to be added before the other data layers (otherwise it
    will cover them)
4.  If we are using these tiles, it is especially important to set the
    map CRS to Web Mercator: `coord_sf(crs = "epsg:3857")`
5.  Change the CRS of data with `project(<data>, "epsg:4326")`
6.  To add north arrows and scale bars, use `annotation_north_arrow()`
    and `annotation_scale()`

Prompt: Try mapping the built-up area. The file is

```{r}
file.path(spatial_dir, "cumilla_wsf_4326.tif")
```

### Using `plot_layer()` and layers.yml

To handle more situations, and so that we don't need to build up the
commands for each plot every time, we have a custom function
`plot_layer()`. To do so we need to set some universal variables. These
are all defined in the top of `R/maps.R`:

```{r}
# Set static map visualization parameters
layer_alpha <- 0.7
map_width <- 6.9 # width of map (excluding legend) in inches
map_height <- 5.9 # height of map in inches
aspect_ratio <- map_width / map_height

# Define map extent and zoom level
static_map_bounds <- aspect_buffer(aoi, aspect_ratio, buffer_percent = 0.05)
  # static_map_bounds makes sure all maps for all cities have the same dimensions
zoom <- round(15.6 + -0.00015 * (sqrt(expanse(aoi))))
tiles <- annotation_map_tile(type = "cartolight", zoom = zoom, progress = "none")
```

You can change the dimensions of all maps by changing `map_width` and
`map_height`.

To plot the AOI, we just need to set `aoi_only = TRUE` and
`plot_aoi = TRUE`. (We can also include ward lines, if we have data for
these, by replacing `plot_aoi = TRUE` with `plot_wards = TRUE`. We can
follow up on how to include wards later.)

```{r}
plot_layer(aoi_only = TRUE, plot_aoi = TRUE)
```

To make the population map, we just need to provide the data and the key
"population".

```{r}
plot_layer(data = pop, yaml_key = "population")
```

What's going on here? `plot_layer()` finds parameters for how to style
the data in `source/layers.yml`, where there is a *dictionary* for
population called "population". It looks like this:

```         
population:
  fuzzy_string: population.*.tif$
  title: 'Population density'
  subtitle: 'Number of persons per 10,000 m^2^'
  palette: ['#ECEB72', '#8E3933']
  bins: 5
```

and tells `plot_layer()` to make a color gradient from yellow to brown
with 5 bins and to use "Population density\nNumber of persons per 10,000
m^2^" as the title. By changing these values, we can change how the
population map is plotted.

We can do the same for built-up extent, which uses the YAML key "wsf"
for World Settlement Footprint.

```{r}
# layer.yml also contains a string to use for reading the relevant file in the spatial directory
wsf_data <- fuzzy_read(spatial_dir, fuzzy_string = layer_params$wsf$fuzzy_string)
plot_layer(data = wsf_data, yaml_key = "wsf")
```

The YAML dictionary for this looks like

```         
wsf:
  fuzzy_string: wsf_4326.tif$
  title: 'Year of urban expansion'
  breaks:  [1985,           1986,        1996,        2006,       2015]
  labels:  ['Before 1986', '1986-1995', '1996-2005', '2006-2015']
  palette: ['#FBF9CD',     '#E3B854',   '#BF4A3B',   '#25110B']
```

Here, we implicitly state the number of bins as 4 with the number of
labels, and we assign a color to each label. The breaks tell `ggplot`
how to divide up the data. Note that we have one more break than label,
because we need to include a closing break (2015). (When the data is
non-continuous, the number of labels and breaks will match. In this
case, the breaks don't divide the data but say which discrete value
should associate with each label.)

### R/maps.R

With these details defined in `layers.yml`, we can loop through all of
the layers for which we have data. That's what this bit of R/maps.R is
doing.

```{r, eval = FALSE}
# Create an empty list for the plots to be saved to
plots <- list()

# Take all of layer names from layer_params as well as the fuzzy file names
unlist(lapply(layer_params, \(x) x$fuzzy_string)) %>%
  # Loop through the fuzzy file names and the layer names
  map2(names(.), \(fuzzy_string, yaml_key) {
    tryCatch({
      # For each file and layer name, read the data and make the plot
      data <- fuzzy_read(spatial_dir, fuzzy_string) %>% vectorize_if_coarse()
      plot <- plot_layer(
        data = data, yaml_key = yaml_key,
        plot_aoi = is.null(wards), plot_wards = !is.null(wards))
      # Save the plot the the plots list
      plots[[yaml_key]] <<- plot
      message(paste("Success:", yaml_key))
    },
    # If there's an error, move on to the next layer!
    error = \(e) warning(glue("Error on {yaml_key}: {e}")))
  })
```

Most of the plots are made this way, but a few require a little more
work.

1.  Some of the maps require a final bit of pre-processing. This is done
    with `source("R/pre-mapping.R")`.
2.  Some maps require multiple layers, such a flooding probability on
    top of population. For this we set `baseplot` to the desired lower
    layer in `plot_layer()`.

```{r}
# Create empty list to hold all of the maps
plots <- list()

# Save population plot to the list
plots$population <- plot_layer(data = pop, yaml_key = "population")
plots$population

fluvial <- fuzzy_read(spatial_dir, "fluvial_2020.tif$")

plot_layer(data = fluvial, yaml_key = "fluvial")

fluvial <- fuzzy_read(spatial_dir, "fluvial_2020.tif$")
plot_layer(data = fluvial, yaml_key = "fluvial", baseplot = plots$population)
```

While there are more steps hidden inside `plot_layer()`, this is
essentially just adding another `geom_spatraster()` on top of the first,
as in the following:

```{r}
ggplot() +
  geom_spatraster(data = pop) +
  scale_fill_stepsn(
    colors = layer_params$population$palette,
    na.value = "transparent"
  ) +
  labs(fill = "Population density") +
  ggnewscale::new_scale_fill() +
  geom_spatraster(data = fluvial) +
    scale_fill_stepsn(
    colors = layer_params$fluvial$palette,
    na.value = "transparent"
  ) +
  geom_spatvector(data = aoi, color = "black", fill = NA)
```

In `R/maps.R` the multi-layering of flood maps happens by sourcing
`R/map-flooding.R`.

Finally, with all the maps saved to the `plots` list, we save all of the
plots

```{r}
# Save plots -------------------------------------------------------------------
plots %>% walk2(names(.), \(plot, name) {
  save_plot(plot, filename = glue("{name}.png"), directory = styled_maps_dir)
})
```

There's a lot that's confusing in this (see `%>%`, `walk2`, `.`), but
essentially it's doing the following for each map in `plots`:

```{r}
save_plot(plots$population, filename = "population-tutorial.png", directory = styled_maps_dir)
```

We could also use `ggsave()` like we did for the charts, but
`save_plot()` is a custom function that ensures the legend is the same
width for all maps. (The default legend width is 1/3 of the map width:
that is the map is 3 parts wide and the legend is 1 part wide. You can
change this by adding `rel_widths = c(X, Y)` to `save_plot()` but note
this hasn't been thoroughly tested; the title line length does not
automatically adapt.)

```{r}
save_plot(
  plots$population, filename = "population-tutorial-new-width.png", directory = styled_maps_dir,
  rel_widths = c(5, 1))
```

#### Takeaways

Ways to modify maps 1. Edit colors, break points, and titles in
`source/layers.yml`. 2. Save as PDFs rather than PNGs by changing "png"
in the `save_plot()` call. 3. Stack layers by using `baseplot` in
`plot_layer()`. 4. Change map dimensions by changing the `map_width` and
`map_height` at the top of `R/maps.R`. This might be helpful for cities
with very wide or tall AOIs. If doing so, it is important to run the
lines defining `aspect_ratio`, `static_map_bounds`, `zoom` and `tiles`
(and easiest to just run all or `R/maps.R`. 5. To some degree, you can
edit an existing map (e.g., `plots$population`) by adding ggplot
functions to override those already included. (This will not work as
well on multi-layer maps.)

Process 1. Final pre-processing happens in `R/pre-mapping.R` 2.
`R/maps.R` uses custom functions defined in `R/fns.R` 3. Most maps are
made by iterating through all of the layers in `source/layers.yml` 4. We
use `save_plot()` instead of `ggsave()` so that all maps have the same
legend width

### Prompts

1.  Try to modify the WSF map by showing only 3 bins instead and using
    different colors.
2.  Layer roads on top of the built-up area map
